---
title: "p8105_hw3_gw2442"
output: github_document
date: "2022-10-07"
---
```{r}
library(tidyverse)
```

## Problem 1 

```{r}
library(p8105.datasets)
data("instacart")

instacart %>%
  group_by(aisle_id) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs))
```
There are 134 aisle. The aisle with the most items ordered from were:

 * Aisle 83 with 150,609 items
 * Aisle 24 with 150,473 items
 * Aisle 123 with 78,493 items

```{r}
instacart %>%
  group_by(aisle_id) %>%
  mutate(
    n_obs = n()
  ) %>%
  filter(n_obs > 10000) %>%
ggplot(aes(x = aisle_id, y = n_obs)) +
  geom_point()
```

```{r}
instacart %>%
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(n_obs = n()) 

```

## Problem 2

Loading, tidying, and wrangling the dataset:
```{r}
accel_data = read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday")) %>%
  select(week, day_id, day, day_type, everything()) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_number",
    names_prefix = "activity_",
    names_transform = list(activity_number = as.numeric),
    values_to = "activity_count_min"
  ) 

skimr::skim(accel_data)
```
The dataset consists of the following variables: week, day_id, day, day_type, activity_number, and activity_count_min. All variables are numeric variables except for day and day_type. The dataset is made up of 50400 rows and 6 columns. 

```{r}
accel_data %>%
  group_by(day_id) %>%
  mutate(
    activity_total = sum(activity_count_min)
  ) %>%
  summarise(day, activity_total) %>%
  distinct %>%
  print(n = 35)
```
From this table, there are no obvious trends apparent. However, two of the Saturdays (day_id = 24, 31) both have a total activity of 1440 minutes. 

```{r}
accel_data %>%
  mutate(
    week = as.character(week)
  ) %>%
  ggplot(aes(x = activity_number, y = activity_count_min, color = week)) + 
  geom_point(alpha = 0.5) +
  labs(
    title = "24 Hour Activity Time Plot",
    x = "Minutes",
    y = "Activity Count"
  )
```
Activity seems to spike around minute 650 and 1250 for weeks 1, 2, 3, and 4. Weeks 1 and 2 also demonstrate a spike in activity during minute 1000. For week 5, activity seems to spike around minute 400 and 1250. Based on this graph, the 63-year old male with BMI 25 demonstrates a rather consistent daily routine for the 5 recorded weeks. 


## Problem 3 

```{r}
library(p8105.datasets)
data("ny_noaa")

problem_3 = ny_noaa %>%
  janitor::clean_names() %>%
  separate(date, c("year", "month", "day"), sep = "-") %>%
  mutate(
    prcp = prcp*10
  )


```









